Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-05 chapters/LBM_technique.tex
Changes 1. Rewritten literature review and introduction to combine both 2. Beginning to rewrite methods section 3. Finished rewrite of aim 1 - magnetic fields 4. Finished adding structure characterization

@@

$Re = \frac{\rho u l}{\eta}$ which defines the ratio between inertial and viscous forces. Another physical phenomenon being described is non-ideal mixing. The free energy of mixing
is defined as $\Delta G_m = \Delta H_m - T \Delta S_m$. In an ideal solution, $\Delta H_m = 0$. An expression for $\Delta S_m$ for a binary mixture is obtained from statistical mechanical 
approximations of the configurations that binary mixtures can take, resulting in

@@

The Shan-Chen pseudopotential
model is used to simulate the Cahn-Hilliard equation, defined below

\begin{equation}
    \begin{split}
    \frac{\partial\phi}{\partial t}+\nabla\cdot\left(\phi\vec{u}\right) &= \nabla \cdot \left( \Gamma  \nabla\phi \right)
    \end{split}
\end{equation}

where $c^k$, $M^k$ and $u^k$ is the concentration, mobility and velocity of species $k$ respectively while $\mu$ is the free energy of system. 
\cite{shan_lattice_1993, shan_simulation_1994, he_lattice_1997, he_discrete_1998}

A computational model designed to simulate soft matter systems 
at the mesoscale must accurately represent both hydrodynamics and non-ideal mixing in order to capture the essential physics of emulsions, colloids, and phase-separating fluids. 
A hydrodynamic model must be able to recover mass and momentum conservation of a incompressible fluid

Modelling multicomponent dynamics requires incorporating a free energy model between fluids, allowing for the implementation of
chemical potential driven mass transfer and surface tension. Typically this is done using a Cahn-Hilliard model based on a
square gradient free energy functional. This captures the demixing characteristics of phase separating fluids and controls interface
characteristics, providing surface tension and interface width characteristics. This model has been used in past investigations looking 
into spinodal decomposition, domain growth and coarsening.

To fully capture mesoscale dynamics, the model must also resolve the coupling between hydrodynamic flow and composition gradients. This coupling 
is essential for simulating phenomena such as Marangoni flows, interfacial instabilities, and anisotropic stress distributions that arise from 
non-uniform compositions or curvature-dependent surface tension. Moreover, the model must support anisotropic or time-dependent external fields 
(e.g., magnetic, electric, or shear) that influence mixing behavior. Computational frameworks that can evolve velocity, pressure, and composition 
fields simultaneously—while ensuring thermodynamic consistency and mechanical stability—are critical for accurately describing the emergent structures 
and transport properties in non-ideal soft matter systems.

Given the multiscale modeling necessary to correctly simulate the physical system, one approach suggested to tackle this problem is the multicomponent
Lattice Boltzmann Method. This technique bridges the length and time scales necessary in soft matter modeling through modeling the evolution of a particle
distribution function. Hydrodynamic parameters are recovered through calculating the moments of the particle distribution function. One of the advantages of
the Lattice Boltzmann Method its parallelize nature and ease of adding multiple physical phenomena. In the following sections, the implementation of the LBM
and alternatives to the LBM will be presented.

>>>>

\section{Hydrodynamics}

@@

$Re = \frac{\rho u l}{\eta}$ which defines the ratio between inertial and viscous forces.

\section{Phase separation}

Another physical phenomenon being described is non-ideal mixing. The free energy of mixing
is defined as $\Delta G_m = \Delta H_m - T \Delta S_m$. In an ideal solution, $\Delta H_m = 0$. An expression for $\Delta S_m$ for a binary mixture is obtained from statistical mechanical 
approximations of the configurations that binary mixtures can take, resulting in

@@

\begin{equation}
    \begin{split}
    \frac{\partial\phi}{\partial t} = \nabla \cdot \left( \Gamma  \nabla\mu_\phi \right)
    \end{split}
\end{equation}

\section{Particle dynamics}

\section{Magnetic fields and particles}

% The Shan-Chen pseudopotential
% model is used to simulate the Cahn-Hilliard equation, defined below

% A computational model designed to simulate soft matter systems 
% at the mesoscale must accurately represent both hydrodynamics and non-ideal mixing in order to capture the essential physics of emulsions, colloids, and phase-separating fluids. 
% A hydrodynamic model must be able to recover mass and momentum conservation of a incompressible fluid

% Modelling multicomponent dynamics requires incorporating a free energy model between fluids, allowing for the implementation of
% chemical potential driven mass transfer and surface tension. Typically this is done using a Cahn-Hilliard model based on a
% square gradient free energy functional. This captures the demixing characteristics of phase separating fluids and controls interface
% characteristics, providing surface tension and interface width characteristics. This model has been used in past investigations looking 
% into spinodal decomposition, domain growth and coarsening.

% To fully capture mesoscale dynamics, the model must also resolve the coupling between hydrodynamic flow and composition gradients. This coupling 
% is essential for simulating phenomena such as Marangoni flows, interfacial instabilities, and anisotropic stress distributions that arise from 
% non-uniform compositions or curvature-dependent surface tension. Moreover, the model must support anisotropic or time-dependent external fields 
% (e.g., magnetic, electric, or shear) that influence mixing behavior. Computational frameworks that can evolve velocity, pressure, and composition 
% fields simultaneously—while ensuring thermodynamic consistency and mechanical stability—are critical for accurately describing the emergent structures 
% and transport properties in non-ideal soft matter systems.

% Given the multiscale modeling necessary to correctly simulate the physical system, one approach suggested to tackle this problem is the multicomponent
% Lattice Boltzmann Method. This technique bridges the length and time scales necessary in soft matter modeling through modeling the evolution of a particle
% distribution function. Hydrodynamic parameters are recovered through calculating the moments of the particle distribution function. One of the advantages of
% the Lattice Boltzmann Method its parallelize nature and ease of adding multiple physical phenomena. In the following sections, the implementation of the LBM
% and alternatives to the LBM will be presented.
